\myTitle

### 2014-12-02 tex/main-body-text.tex
Changed results, with new figs, cleaned figure directory out

\section*{Comparison of {\sc structure} and {\sc flock} results}
Data sets with known $K$ were simulated
(code provided on GitHub repository https://github.com/eriqande/flock-comment)
and evaluated in the programs {\sc flock} and {\sc structure} 
for the known value.The first dataset simulated was composed of five populations of 
sizes 500, 500, 800, 1100, and 1100 comprised of individuals 
with 15 loci and an average pairwise $F_{ST} = 0.03$ between populations. 
The second and third dataset were simulated with population sizes
250, 250, 400, 550, and 550 and average pairwise $F_{ST} = 0.02 and 0.03$
respectively. {\sc flock} was run nine times using the default setting for all 
run parameters (initial partition = random, number of iterations = 20, number of runs = 50, 
LLOD threshold = 0) for the known value of $K$. The no-admixture, 
uncorrelated-allele-frequency model implemented in {\sc structure} was run nine times with 
a 5,000 sweep burn-in followed by a 20,000 sweep sample from the posterior. 
% Looking back some of the runs were under different conditions, not huge differences. 
% But 1000 burn-in and 5,000 sweep look about the same
All runs were performed on a JNCS D685 with an i7 3.40GHz processor and 8GB
of RAM running 64-bit Windows 7.

@@

\citet{Duc&Tur2012} describe a method for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must specify both the number of runs (different initial allocations of individuals 
to each cluster) as well as the number of iterations. (In each iteration every individual has the opportunity to be reallocated 
to a different cluster.) At the conclusion of each run, the log-likelihood value for membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is calculated for each individual. {\sc flock} identifies runs that 
have converged to the same solution by detecting that they have the same mean LLOD score, and, in 
{\sc flock} parlance, a series of runs that have converged to the same solution are termed a ``plateau.'' 
The plateau record is a list of how many runs converged to the same allocation
of individuals. 
\citet{Duc&Tur2012} developed two stopping rules that they use to estimate $K$.\citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau of
$\geq 6$ runs is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a single plateau of $\geq 6$ runs is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau of $\geq 6$ runs, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus of $\geq 6$ runs are observed, $K$
is declared {\em undecided}.

It is now common practice to use the approximate marginal likelihood ($\ln P(D)$) from {\sc structure}
or a derived quantity such as $\Delta K$ \citep{Evannoetal2005} to try to estimate $K$.
While we take the stance that estimates of $K$ made with {\em any} unsupervised clustering algorithm
from data on real (non-idealized) populations should always be interpreted and used
cautiously, it is nonetheless instructive to compare \citet{Duc&Tur2012}'s method for
estimating $K$ to results obtained using $\ln P(D)$ and $\Delta K$, and we do so, below.

\subsection*{Results} 
Run times for both programs were relatively short, albeit in the case of {\sc flock} 
non-trivial. {\sc flock} averaged 67.3 minutes
for each run of the first simulated dataset, while each group of 
9 iterations for {\sc structure} averaged 18 minutes. CALCULATE
TIME FOR OTHER SIMULATIONS

{\sc flock} and {\sc structure} provide nearly identical results in 
terms of how individuals are allocated among runs for 
all simulations (Fig.~\ref{fig:Sims}). As $F_{ST}$ decreases 
between populations the solutions that {\sc flock} are more variable. 
The average Euclidean distance between the known $q_i$ value and 
the infered value from each program increases as the genetic 
differentiation among populations decreases. {\sc flock} had an 
smaller mean Euclidean distance over all runs for simulations 
1 and 2, but had much larger variation in simulation 3 (Fig.~\ref{fig:ED}).
%calculate Euclidean distance between flock and structure values.

There was very little variation in the 'best-run' returned by {\sc flock}. 
From the nine different sets of runs for the first simulated dataset, 
three groups of runs ended with 
identical clusters (runs 2,8 \& 9, runs 1 \& 3, and runs 4 \& 7). 
Of the other 31 pairwise comparisions that were not identical
15 differed by the placement of 1 indvidual, 11 differed by the placement
of 2 individuals, and 5 differed by the placement of 3 individuals.

Despite the similar allocation of individuals into $K$ clusters, the 
plateau analysis to estimate $K$ in {\sc flock} did not support $K = 5$. 
For the first simulated dataset plateaus of varioius lengths were observed, 
but none greater than 5 while no plateaus were observed for either second or 
third simulated datasets.

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{images/Figures-Pat/Simulations_Fig1.pdf}   % Use this if we want it color (I'll make greyscale later)  
    \caption{Plots of $q_i$ values from {\sc structure} and the normalized likelihood values 
of {\sc flock} made with the program {\sc distruct}. Each horizontal plot represents one of nine individual runs of the program {\sc structure}, or the `best run' from a series of nine runs in {\sc flock} at $K=5$. Within each plot 
$q_i$ values from {\sc structure} and the normalized likelihood values 
from {\sc flock}  are represented by a 
%gray-scale
vertical bar. Simulation 1 was composed of populations of 
size 500, 500, 800, 1100, and 1100 with an average pairwise $F_{ST} = 0.03$ between populations. 
The second and third dataset were simulated with population sizes
250, 250, 400, 550, and 550 and average pairwise $F_{ST} =$ 0.02 and 0.03
respectively.}
    \label{fig:Sims}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
    \includegraphics[width=0.5\textwidth]{images/Figures-Pat/EucliDist.pdf}   % Use this if we want it color (I'll make greyscale later)  
    \caption{Eucledian distance between true $q_i$ value and inferred value from {\sc structure} and {\sc flock} .}
    \label{fig:ED}
\end{center}
\end{figure*}

>>>>

\section*{Comparison of {\sc structure} and {\sc flock} results}
Data sets were simulated (code provided on GitHub repository 
https://github.com/eriqande/flock-comment)
and evaluated in the programs {\sc flock} and {\sc structure} 
for the known value ($K=5$). Each dataset was composed of five populations of 
sizes 250, 250, 400, 550, and 550 comprised of individuals 
with 15 loci and an average pairwise $F_{ST} = 0.03, 0.02, 0.01, and 0.005$ among
populations. {\sc flock} was run nine times at $K=5$ with the default setting for all 
run parameters (initial partition = random, number of iterations = 20, number of runs = 50, 
LLOD threshold = 0). The default model in {\sc structure}, the no-admixture and correlated
allele frequency (NA-C) model, the no-admixture and non-correlated allle frequency (NA-NC) model, 
and the no-admixture and non-correlated allele frequency (NA-NC) model 
were run nine times with a 5,000 sweep burn-in followed by a 20,000 sweep sample 
from the posterior. All runs were performed on a JNCS D685 with an i7 3.40GHz processor and 8GB
of RAM running 64-bit Windows 7.

@@

\citet{Duc&Tur2012} describe a method for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must specify both the number of runs (different initial allocations of individuals 
to each cluster) as well as the number of iterations. (In each iteration every individual has the opportunity to be reallocated 
to a different cluster.) At the conclusion of each run, the log-likelihood value for membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is calculated for each individual. {\sc flock} identifies runs that 
have converged to the same solution by detecting that they have the same mean LLOD score, and, in 
{\sc flock} parlance, a series of runs that have converged to the same solution are termed a ``plateau.'' 
The plateau record is a list of how many runs converged to the same allocation
of individuals. 
\citet{Duc&Tur2012} developed two stopping rules that they use to estimate $K$. 
\citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau of
$\geq 6$ runs is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a single plateau of $\geq 6$ runs is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau of $\geq 6$ runs, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus of $\geq 6$ runs are observed, $K$
is declared {\em undecided}.

%This section may get cut depending on how much we want to talk about estimating K.
It is now common practice to use the approximate marginal likelihood ($\ln P(D)$) from {\sc structure}
or a derived quantity such as $\Delta K$ \citep{Evannoetal2005} to try to estimate $K$.
While we take the stance that estimates of $K$ made with {\em any} unsupervised clustering algorithm
from data on real (non-idealized) populations should always be interpreted and used
cautiously, it is nonetheless instructive to compare \citet{Duc&Tur2012}'s method for
estimating $K$ to results obtained using $\ln P(D)$ and $\Delta K$, and we do so, below.

\subsection*{Results} 
Run times for both programs were relatively short, albeit in the case of {\sc flock} 
non-trivial. {\sc flock} averaged 32.8 minutes
for each run of the first simulated dataset, while each group of 
9 iterations for {\sc structure} averaged 14.7, 15.3 and 18.7 minutes
for the NA-NC, NA-C, and A-C models respectively.

{\sc flock} and {\sc structure} provide nearly identical results in 
terms of how individuals are allocated among runs for 
all simulations (Fig.~\ref{fig:Sims}). As $F_{ST}$ decreases 
between populations the solutions that {\sc flock} found were more variable.
Of the nine runs of {\sc flock} for the first simulated dataset, between 0 and 22
differences between the final arrangement of indivuals within the ($K=5$)
clusters. Of the nine runs, four pairs of runs found the 
same solution (runs 1 \& 2, 5 \& 6, 3 \& 7, and 4\&9). Run 8, which 
had no plateaus, was the most different with an average of 19.86 differences.
For the third simulated dataset, the number 
of differences ranged from 482 to 911.  {\sc structure} showed less variation for 
each set of simulations. Solutions for the first simulated dataset differed by the
placement of between 0 and 4 individuals and between 5 and 17 for the third 
dataset.
%Create a figure comparing all of the values. All the text is somewhat hard to read

Plateau lengths greater than 0 were only obsered for the first simulated dataset. For each
of the nine runs of {\sc flock} the plateau lengths varied from 0 to 4. In no analysis 
did we observe a plateau lenght greater than 6. Despite the similar allocation of individuals into $K$ clusters, the 
plateau analysis to estimate $K$ in {\sc flock} did not support $K = 5$.

but none greater than 5 while no plateaus were observed for either second or 
third simulated datasets.

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{images/Figures-Pat/Fig_AllSim.pdf}   % Use this if we want it color (I'll make greyscale later)  
    \caption{Plots of $q_i$ values from {\sc structure} and the normalized likelihood values 
of {\sc flock} made with the program {\sc distruct}. Each horizontal plot represents one of nine individual runs of the program {\sc structure}, or the `best run' from a series of nine runs in {\sc flock} at $K=5$. Within each plot 
$q_i$ values from {\sc structure} and the normalized likelihood values 
from {\sc flock}  are represented by a 
%gray-scale
vertical bar. Each dataset is comprised of five populations of sizes 550, 550, 400, 250, and 250 individuals. The 
average pairwise $F_{ST}$ among populations was 0.03, 0.02, 0.01, and 0.005 for simulations 1, 2, 3, and 4 respecitively. 
}
    \label{fig:Sims}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
    \includegraphics[width=0.5\textwidth]{images/Figures-Pat/Loss.pdf}
    \caption{Comparision of zero-One loss for the admixture and correlated allele frequency (A-C) model, 
the no-admixture and correlated allele frequency (NA-C) model, the no-admixture and non-correlated 
allle frequency (NA-NC) model, the no-admixture and non-correlated allele frequency (NA-NC) 
model in {\sc structure} and the {\sc flock} model for each of the four simulated datasets.}
    \label{fig:Loss}
\end{center}
\end{figure*}
